---
title: "206 Assignment 3 Task 3"
author: "Alex Gaynor"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Exploring Urchins Data

Here I explore data from the coastline near Santa Barbara about sea urchins.

```{r, message = FALSE, warning = FALSE}
urchins <- read.csv(here("data", "urchins.csv")) # read in data
urchins_clean <- clean_names(urchins) # clean names
```

### Exploring Size Comparisons With Plots

A histogram faceted by urchin type.

```{r, message = FALSE, warning = FALSE}
ggplot(urchins_clean, aes(x = size, fill = common_name)) +
  geom_histogram() +
  facet_wrap(~ common_name) +
  xlab("Common Name") +
  ylab("Size") +
  theme(legend.position = "none")
```

A boxplot of urchin sizes by common name.

```{r, message = FALSE, warning = FALSE}
ggplot(urchins_clean, aes(x = common_name, y = size, fill = common_name)) +
  geom_boxplot(legend = FALSE) +
  xlab("Common Name") +
  ylab("Size") +
  theme(legend.position = "none")
```

Creating a table with measures of central tendency

```{r, message = FALSE, warning = FALSE}
# summarize summary statistics
urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean_size_cm = mean(size),
            median_size_cm = median(size),
            sd_size_cm = sd(size),
            count = n())
# place summary stats in a table with nice titles
urchins_summary %>%
  kable(col.names = c("Common Name",
                    "Mean Size (cm)",
                    "Median Size (cm)",
                    "Standard Deviation",
                    "Sample Size (n)")) %>%
  kable_styling(full_width = FALSE, "hover")
```

The Red Urchins are larger than the Purple Urchins on average, by about 3cm. The standard deviation of Red Urchins is larger which means their measurements have a higher variance relative to the mean. This means the range of size for Red Urchins is larger.

## Explore Purple Urchins


```{r, message = FALSE, warning = FALSE}
# Organize the Purple Urchins Data
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>%
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))
# Place data in table
purple_urchins_table <- purple_urchins %>% 
  group_by(year) %>%
  count(site_full)
#Create graph of data
ggplot(purple_urchins_table, aes(x = year, y = n, group = site_full, colour = site_full)) +
  geom_line(size = 1, alpha = 0.6) +
  scale_x_continuous(breaks = c(2008, 2012, 2016, 2020))
  
```

